---
title: "Course 8 Project"
author: "Letitia Downen"
date: "August 22, 2017"
output: 
  html_document: 
    keep_md: yes
---

## Summary

Using Devices such as *Jawbone up*, *Nike Fuelband*, and *Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively.  These types of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in behavior, or because they are teck geeks. In this project, we will use accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

This project will attempt find a model to predict the manner in which they did the exercise (A,B,C,D,E)

## Data Analysis/Set Creation

```{r}
library(caret)
training<-read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",na.strings = c("NA","#DIV/0!",""))
                    
testing<-read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",na.strings = c("NA","#DIV/0!",""))

```
Remove NA columns and Non informative columns(columns 1 to 7)

```{r}
training<-training[,colSums(is.na(training))==0]
testing<-testing[,colSums(is.na(training))==0]

training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]
```

create training and validation sets from the full training set.
```{r}
set.seed(35)
inTrain<-createDataPartition(training$classe,p=.7)[[1]]
mytrain<-training[inTrain,]
myvalid<-training[-inTrain,]
```
## Model Selection
Set the control for cross validation.  We will use this for each model, so it is useful to set it ahead of time.
```{r}
control<-trainControl(method="cv",number=5)
```
I. Decision Trees
```{r}
set.seed(98)
rpartmod<-train(as.factor(classe)~.,data=mytrain,method="rpart")

rpartpred<-predict(rpartmod,newdata = myvalid)

confusionMatrix(rpartpred,myvalid$classe)$overall["Accuracy"]

```

II. Random Forests with Preprocessing

```{r}
set.seed(98)
rfmod<-train(classe~.,data=mytrain,method="rf",trControl=control,preProcess="pca")
rfpred<-predict(rfmod,newdata=myvalid)
confusionMatrix(rfpred,myvalid$classe)$overall["Accuracy"]

```
II. Boosting with Trees(gbm)

```{r}
set.seed(98)
gbmmod<-train(classe~.,data=mytrain,method="gbm",trControl=trainControl(method="cv",number=5),verbose=FALSE)

gbmpred<-predict(gbmmod,newdata=myvalid)
confusionMatrix(gbmpred,myvalid$classe)$overall["Accuracy"]


```

The method of using Random Forests has a high level of accuracy and therefore, I will not combine predictors/models.  

## Predictions

We will use the random forest model to predict the manner in which 20 participants exercised.

```{r}
predict(rfmod,newdat=testing)
```

